Decoding the Amplitude-Modulated Part the DCF77 Longwave Time Signal by using the Goertzel Algorithm

### 2024-08-29 paper/main.tex
Demo implementation

\subsection{Results}
Multiple simulations with different values for the SNR starting from \SIrange{5}{-7}{\deci\bel} were executed to verify and evaluate the receiver's capabilities to reconstruct the DCF77's time code signal in different receiving conditions.
From these simulations, the most relevant results occured in the case where the SNR was equal to \SI{-5}{\deci\bel}, because this was the lowest SNR where the reciver was still able to
retrieve back the time code signal without errors.
The frequency domain plots of the two FFTs, shown in Fig. \ref{fig:sim_fd_m5dB}, verify the DCF77 signal's aliases to show up exactly at \SI{\pm 5.5}{\kilo\hertz}.
From the plot showing the DCF77 signal after passing the AWGN-channel, it can be seen that the noise floor is located at approximately \SI{-50}{\deci\bel} while the peaks
of the DCF77 signaö have a value of approximately \SI{-8}{\deci\bel}.
The difference of \SI{42}{\deci\bel} should not be confused with the SNR, as the power of the noise added by the AWGN-Channel is spread evenly over all frequencies.
The correct SNR was verified to be exactly \SI{-5}{\deci\bel} by using MATLAB's built-in command "snr()" to calculate the SNR by comparing the before and after adding noise.
\begin{figure}[!htbp]
    \centerline{\includegraphics[width=0.4\textwidth]{img/sim_fd_m5dB.eps}}
    \caption{Plots of the frequency domain at a SNR of \SI{-5}{\deci\bel}}
    \label{fig:sim_fd_m5dB}
\end{figure}
\FloatBarrier
The time domain plots of Fig. \ref{fig:sim_td_m5dB} show the creation of the DCF77 signal and the internal operation of the receiver.
The first plot shows the time code signal that should be retrieved back by the receiver to allow a visual verification that the DCF77 signal is modulated retrieved back correctly.
It can be seen that each symbol has a length of \SI{1}{\second} with a gap of either \SI{200}{\milli\second} or \SI{100}{\milli\second} at the beginning.
From the second plot, the amplitude can be seen to switch between $0.15$ and $1$, representing the changes between \SI{15}{\percent} and \SI{100}{\percent}.
The third plot shows the DCF77 signal with added noise and an average amplitude of approximately $3.6$ with the orignal waveform not being visible anymore.
The fourth plot shows the results of the Goertzel algorithm segments together wit the value of the calculated threshold.
The operation of the Goertzel algorithm can be verified, as the calculated magnitude of the frequency bin is clearly following the waveform of the time code signal.
The value propagation of the threshold is as slow as expected with the value only experiencing minor drops during the gaps of the signal.
A value where the detector starts to take valid decisions is reached after approximately \SI{5}{\second}.
The threshold then approaches asymptotically the value where it is exactly located in the middle between the area of values representing an apmlitude of \SI{15}{\percent} and
the area of values representing \SI{100}{\percent}.
As the value of the EMA is proportional to the threshold, the plot is left out.
In the fifth plot, the retrieved time code signal is visualized.
It is obvious that the signal contains errors until the threshold has reached a level where valid decisions are taken by the detector.
After \SI{5}{\second}, when the threshold reached this value, no more errors are visible.
\par
All simulation runs with a smaller SNR contained an increasing amount of errors, caused by the noise to raise or drop single points above or below the threshold, placing them in the
area where they were detected to correspond to the opposite symbol.
This effect can be imagined by thinking of the height of the area where the results are distributed by the noise is increased until both areas touch the threshold and start to overlap.
Single points ending up in the other area are then interpreted as the wrong symbol.

@@

@@

\Urlmuskip=0mu plus 1mu
\begin{thebibliography}{00}
\bibitem{b1} -, ``DCF77,'' 2024, Wikipedia entry. [Online]. Available: \url{https://de.wikipedia.org/wiki/DCF77}
\bibitem{b2} D. Priester and P. Hetzel and A. Bauch, ``Zeit- und Normalfrequenzverbreitung mit DCF77,'' PTB Mitteilungen, 2004, [Online]. Available: \url{https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung_4/4.4_zeit_und_frequenz/4.42/dcf77.pdf}
\bibitem{b3} D. Priester and P. Hetzel and A. Bauch, ``Zeit- und Normalfrequenzverbreitung mit DCF77: 1959-2009 und darüber hinaus,'' PTB Mitteilungen, 2009, [Online]. Available: \url{https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung_4/4.4_zeit_und_frequenz/pdf/2009_Bauch_PTBM__DCF77.pdf}
\bibitem{b4} P. Hetzel, ``Time dissemination via the LF transmitter DCF77 using a pseudo-random phase-shift keying of the carrier,'' PTB Braunschweig, 1988, [Online]. Available: \url{https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung_4/4.4_zeit_und_frequenz/pdf/5_1988_Hetzel_-_Proc_EFTF_88.pdf}
\bibitem{b5} -, ``DCF77 webpage,'' 2024, Official page of the PTB [Online]. Available: \url{https://www.ptb.de/cms/en/ptb/fachabteilungen/abt4/fb-44/ag-442/dissemination-of-legal-time/dcf77.html}
\bibitem{b6} R. G. Lyons, ``Understanding Digital Signal Processing,'' Third Edition, Prentice Hall, 2010.
\bibitem{b9} J. Hoffmann, ``Sampling bandpass signals,'' 2021, Internet Article. [Online]. Available: \url{https://www.dsprelated.com/showarticle/1407.php}
\bibitem{b10} F. Zaplata and M. Kasal, ``SDR Implementation for DCF77,'', 2013, IEEE Paper. [Online]. Available: \url{https://ieeexplore.ieee.org/document/6530943}
\bibitem{b7} -, ``Goertzel algorithm,'' 2024, Wikipedia entry. [Online]. Available: \url{https://en.wikipedia.org/wiki/Goertzel_algorithm}
\bibitem{b8} K. Banks, ``The Goertzel Algorithm,'' 2002, Internet Article. [Online]. Available: \url{https://courses.cs.washington.edu/courses/cse466/12au/calendar/Goertzel-EETimes.pdf}
\bibitem{b11} R. G. Lyons, ``Setting the 3-dB Cutoff Frequency of an Exponential Averager,'' 2012, Internet Article. [Online]. Available: \url{https://www.dsprelated.com/showarticle/182.php}
\end{thebibliography}
\end{document}

>>>>

\subsection{Results}
Multiple simulations with different values for the SNR starting from \SIrange{5}{-7}{\deci\bel} were executed to verify and evaluate the receiver's capabilities to reconstruct the DCF77's time code signal in different receiving conditions.
From these simulations, the most relevant results occured in the case where the SNR was equal to \SI{-5}{\deci\bel}, because this was the lowest SNR where the reciver was still able to
retrieve back the time code signal without errors.
The frequency domain plots of the two FFTs, shown in Fig. \ref{fig:sim_fd_m5dB}, verify the DCF77 signal's aliases to show up exactly at \SI{\pm 5.5}{\kilo\hertz}.
From the plot showing the DCF77 signal after passing the AWGN-channel, it can be seen that the noise floor is located at approximately \SI{-50}{\deci\bel} while the peaks
of the DCF77 signaö have a value of approximately \SI{-8}{\deci\bel}.
The difference of \SI{42}{\deci\bel} should not be confused with the SNR, as the power of the noise added by the AWGN-Channel is spread evenly over all frequencies.
The correct SNR was verified to be exactly \SI{-5}{\deci\bel} by using MATLAB's built-in command "snr()" to calculate the SNR by comparing the before and after adding noise.
\begin{figure}[!htbp]
    \centerline{\includegraphics[width=0.4\textwidth]{img/sim_fd_m5dB.eps}}
    \caption{Plots of the frequency domain at a SNR of \SI{-5}{\deci\bel}}
    \label{fig:sim_fd_m5dB}
\end{figure}
\FloatBarrier
The time domain plots of Fig. \ref{fig:sim_td_m5dB} show the creation of the DCF77 signal and the internal operation of the receiver.
The first plot shows the time code signal that should be retrieved back by the receiver to allow a visual verification that the DCF77 signal is modulated retrieved back correctly.
It can be seen that each symbol has a length of \SI{1}{\second} with a gap of either \SI{200}{\milli\second} or \SI{100}{\milli\second} at the beginning.
From the second plot, the amplitude can be seen to switch between $0.15$ and $1$, representing the changes between \SI{15}{\percent} and \SI{100}{\percent}.
The third plot shows the DCF77 signal with added noise and an average amplitude of approximately $3.6$ with the orignal waveform not being visible anymore.
The fourth plot shows the results of the Goertzel algorithm segments together wit the value of the calculated threshold.
The operation of the Goertzel algorithm can be verified, as the calculated magnitude of the frequency bin is clearly following the waveform of the time code signal.
The value propagation of the threshold is as slow as expected with the value only experiencing minor drops during the gaps of the signal.
A value where the detector starts to take valid decisions is reached after approximately \SI{5}{\second}.
The threshold then approaches asymptotically the value where it is exactly located in the middle between the area of values representing an apmlitude of \SI{15}{\percent} and
the area of values representing \SI{100}{\percent}.
As the value of the EMA is proportional to the threshold, the plot is left out.
In the fifth plot, the retrieved time code signal is visualized.
It is obvious that the signal contains errors until the threshold has reached a level where valid decisions are taken by the detector.
After \SI{5}{\second}, when the threshold reached this value, no more errors occured.
\par
All simulation runs with a smaller SNR contained an increasing amount of errors, caused by the noise to raise or drop single points above or below the threshold, placing them in the
area where they were detected to correspond to the opposite symbol.
This effect can be imagined by thinking of the height of the area where the results are distributed by the noise is increased until both areas touch the threshold and start to overlap.
Single points ending up in the other area are then interpreted as the wrong symbol.

@@

\section{Demo implementation}
The receiver's structure, shown in Fig. \ref{fig:block_diag_receiver}, was implemented on a STM32F103C8T6, an Arm Cortex-M3 microcontroller, to provide a brief proof of concept.
To for an optical verification, the obtained time code signal is used to turn on and off a Light-Emitting Diode (LED), but this signal
could also be used to decode the DCF77's transmitted time data.
As a tuned antenna together with a pre-amplifier is needed to filter out the DCF77 signal and amplify it to a level where it can be measured by the microcontroller's internal ADC, a minimalistic circuit 
consisting of a tank circuit tuned to \SI{77.5}{\kilo\hertz} and a two-stage transistor amplifier was built based on the circuit shown in \cite{b12}.
The rest of the of the receiver's structure was implemented in software using C and single-precision floating-point for the caluclations.
\par
A timer was set up to create an internal Pulse Width Modulation (PWM) signal to trigger the conversion of the microcontroller's internal ADC at the sampling frequency of \SI{24}{\kilo\hertz}.
The samples are transferred using circular Direct Memory Access (DMA) into a buffer of the size of $480$.
This size was chosen because it is equal to twice the amount of samples that are needed for the analyzation of one segment.
During runtime, the DMA hardware module fills this buffer and two interrupts are executed when the buffer was filled half way and when it is completely full.
Because of the circular setup of the DMA, it fills the buffer continuously and wraps around back to the first address, after reaching the buffer's end.
This allows a continuous operation of the application with a precise sampling frequency while eliminating the need to handle ADC conversions manually. 
As a segment is ready to be analyzed with every interrupt, the operations needed to process the signal processing are executed inside this Interrupt Service Routine (ISR).
The result of each segment after the detector is then used to turn on or off a General Purpose Input Output (GPIO) that is connected to a LED.
\par
Although its rather simplistic implementation, the demo implementation worked right away.
The observed effect of the receiver to adapt to the current receiving conditions by first outputting wrong results because of a
not matching threshold is also noticeable when running the demo application.
After powerup and changes in the environment and therefore during the adaption of the threshold to its new value, the LED shows
arrythmic behavior.
This

The demo implementation worked right away, showing a good sensitivity

@@

\Urlmuskip=0mu plus 1mu
\begin{thebibliography}{00}
\bibitem{b1} -, ``DCF77,'' 2024, Wikipedia entry. [Online]. Available: \url{https://de.wikipedia.org/wiki/DCF77}
\bibitem{b2} D. Priester and P. Hetzel and A. Bauch, ``Zeit- und Normalfrequenzverbreitung mit DCF77,'' PTB Mitteilungen, 2004, [Online]. Available: \url{https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung_4/4.4_zeit_und_frequenz/4.42/dcf77.pdf}
\bibitem{b3} D. Priester and P. Hetzel and A. Bauch, ``Zeit- und Normalfrequenzverbreitung mit DCF77: 1959-2009 und darüber hinaus,'' PTB Mitteilungen, 2009, [Online]. Available: \url{https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung_4/4.4_zeit_und_frequenz/pdf/2009_Bauch_PTBM__DCF77.pdf}
\bibitem{b4} P. Hetzel, ``Time dissemination via the LF transmitter DCF77 using a pseudo-random phase-shift keying of the carrier,'' PTB Braunschweig, 1988, [Online]. Available: \url{https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung_4/4.4_zeit_und_frequenz/pdf/5_1988_Hetzel_-_Proc_EFTF_88.pdf}
\bibitem{b5} -, ``DCF77 webpage,'' 2024, Official page of the PTB [Online]. Available: \url{https://www.ptb.de/cms/en/ptb/fachabteilungen/abt4/fb-44/ag-442/dissemination-of-legal-time/dcf77.html}
\bibitem{b6} R. G. Lyons, ``Understanding Digital Signal Processing,'' Third Edition, Prentice Hall, 2010.
\bibitem{b9} J. Hoffmann, ``Sampling bandpass signals,'' 2021, Internet Article. [Online]. Available: \url{https://www.dsprelated.com/showarticle/1407.php}
\bibitem{b10} F. Zaplata and M. Kasal, ``SDR Implementation for DCF77,'', 2013, IEEE Paper. [Online]. Available: \url{https://ieeexplore.ieee.org/document/6530943}
\bibitem{b7} -, ``Goertzel algorithm,'' 2024, Wikipedia entry. [Online]. Available: \url{https://en.wikipedia.org/wiki/Goertzel_algorithm}
\bibitem{b8} K. Banks, ``The Goertzel Algorithm,'' 2002, Internet Article. [Online]. Available: \url{https://courses.cs.washington.edu/courses/cse466/12au/calendar/Goertzel-EETimes.pdf}
\bibitem{b11} R. G. Lyons, ``Setting the 3-dB Cutoff Frequency of an Exponential Averager,'' 2012, Internet Article. [Online]. Available: \url{https://www.dsprelated.com/showarticle/182.php}
\bibitem{b12} -, ``ATTINY85 als DCF77-Empfänger,'' 2021, Board Entry. [Online]. Available: \url{https://www.mikrocontroller.net/topic/529477}
\end{thebibliography}
\end{document}
